---
title: "report update"
author: "CUI WANG, Yahui"
date: "2025-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df_clean <- df
df_clean[df_clean == ""] <- NA
df_clean$Belief <- as.character(df_clean$Belief)

df_clean$Belief[df_clean$Belief %in% c("Y", "Yes")] <- "Yes"
df_clean$Belief[df_clean$Belief %in% c("N", "No")]  <- "No"

# 
df_clean$Belief[!df_clean$Belief %in% c("Yes", "No")] <- NA

df_clean$Belief <- factor(df_clean$Belief, levels = c("No", "Yes"))
for (v in c("PH", "MH")) {
  df_clean[[v]] <- as.character(df_clean[[v]])
  df_clean[[v]][df_clean[[v]] %in% c("Y", "Yes")] <- "Yes"
  df_clean[[v]][df_clean[[v]] %in% c("N", "No")]  <- "No"
  df_clean[[v]][!df_clean[[v]] %in% c("Yes", "No")] <- NA
  df_clean[[v]] <- factor(df_clean[[v]], levels = c("No", "Yes"))
}
df_clean$Gender <- as.character(df_clean$Gender)

df_clean$Gender[df_clean$Gender == "M"]    <- "Male"
df_clean$Gender[df_clean$Gender == "F"]    <- "Female"
df_clean$Gender[df_clean$Gender == "GD"]   <- "Gender diverse"
df_clean$Gender[df_clean$Gender == "PNTS"] <- "Prefer not to say"

df_clean$Gender <- factor(
  df_clean$Gender,
  levels = c("Male", "Female", "Gender diverse", "Prefer not to say")
)
df_clean$Smoker <- as.character(df_clean$Smoker)

df_clean$Smoker[df_clean$Smoker %in% c("Y", "Yes", "Current")] <- "Yes"
df_clean$Smoker[df_clean$Smoker %in% c("N", "No")] <- "No"

df_clean$Smoker[!df_clean$Smoker %in% c("Yes", "No")] <- NA

df_clean$Smoker <- factor(df_clean$Smoker, levels = c("No", "Yes"))
df_clean$Age[df_clean$Age == 0] <- NA
df_clean$SES5 <- factor(
  df_clean$SES5,
  levels = c("L1", "L2", "L3", "L4", "L5"),
  ordered = TRUE
)
str(df_clean)
summary(df_clean)

table(df_clean$Belief, useNA = "ifany")
table(df_clean$PH, useNA = "ifany")
table(df_clean$MH, useNA = "ifany")
table(df_clean$Gender, useNA = "ifany")

```

```{r}
model_primary <- glm(
  Belief ~ PH + MH,
  data = df_clean,
  family = binomial
)

summary(model_primary)

```

A binomial logistic regression was conducted to examine whether physical or mental health issues were associated with the belief that one’s condition would worsen while waiting for treatment.
Physical health issues were not significantly associated with this belief (β = 0.18, SE = 0.44, p = 0.69). Similarly, mental health issues were not significantly associated with believing that the condition would worsen (β = −0.47, SE = 0.40, p = 0.24).
Overall, the model showed no evidence that the presence of physical or mental health issues influenced beliefs about worsening health while waiting for treatment.


```{r}
prop.table(table(df_clean$PH))
prop.table(table(df_clean$MH))

```


22.2% of participants reported a physical health issue, and 31.8% reported a mental health issue.


```{r}
#Q2
model_PH <- glm(
  PH ~ Age + Gender + SES5 + Smoker,
  data = df_clean,
  family = binomial
)

summary(model_PH)


model_MH <- glm(
  MH ~ Age + Gender + SES5 + Smoker,
  data = df_clean,
  family = binomial
)

summary(model_MH)

# after rejust  prevalence
mean(predict(model_PH, type = "response"), na.rm = TRUE)
mean(predict(model_MH, type = "response"), na.rm = TRUE)


```

A binomial logistic regression was conducted to assess the prevalence of physical health issues in the cohort, adjusting for age, gender, socioeconomic status (SES5), and smoking status.
Age was not significantly associated with the presence of physical health issues (β = 0.01, p = 0.75). There were also no significant differences in physical health issues by gender, including female, gender diverse, or prefer not to say categories (all p > 0.05).

```{r}
#Q3
summary(df_clean$Age)
prop.table(table(df_clean$Gender))
prop.table(table(df_clean$SES5))
prop.table(table(df_clean$PH))
prop.table(table(df_clean$MH))
prop.table(table(df_clean$Smoker))

```

#predominantly middle-aged sample
#Gender distribution was relatively balanced
#This indicates reasonable representation across socioeconomic groups, although slight variation between quintiles was observed.
#Regarding health characteristics, 22.2% of participants reported having a physical health issue, while 31.8% reported a mental health issue. In contrast, a high proportion of participants reported current smoking (85.8%), suggesting that smokers may be over-represented in the sample.